---
title: "Learning Bayesian Data Analysis 1"
output:
  html_document:
    number_sections: yes
    theme: cerulean
    toc: yes
---
# What is This Stuff Called Probability?

## Excercise 4.1

Point of excercises is also to get used to R


Built in library HairEyeColor
Show the data in `HairEyeColor`
```{r}
show(HairEyeColor)
```
The data given is for both male and female. We sum the frequencies across sex.


```{r Sum Freq across Sex}
EyeHairFreq=apply(HairEyeColor, c("Eye","Hair"),sum)
show(EyeHairFreq)
```


Calculate Joint Proportions
```{r Joint Proportions}
EyeHairProp=EyeHairFreq/sum(EyeHairFreq)
show(round(EyeHairProp,2))
```
Marginal proportions across Hair and Sex
```{r Marginal Proportions}
HairFreq=apply(HairEyeColor, c("Hair"),sum)
HairProp=HairFreq/sum(HairFreq)
show(round(HairProp,2))
```
Marginal proportions across Eye and Sex
```{r}
EyeFreq=apply(HairEyeColor, c("Eye"),sum)
EyeProp=EyeFreq/sum(EyeFreq)
show(round(EyeProp,2))
```

Conditional probablity of Hair colours given Blue eyes. We normalise by the marginal probability of Blue eyes by `EyeProp["Blue"]`
```{r}
EyeHairProp["Blue",]/EyeProp["Blue"]
```

Probability of Hair colour given brown eyes
```{r echo=FALSE}
EyeHairProp["Brown",]/EyeProp["Brown"]
```


Probability of Eye colour given brown hair


**NOTE: the 'EyeHairProp[,"Brown"]' gives marginal probabilities along the column**
```{r, echo=TRUE}
EyeHairProp[,"Brown"]/HairProp["Brown"]
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.

## Excercise 4.2 Simulate biased coib with $p(H)=8$

Based on RunningProportion.R 

### Original Program

Specify the total number of flips, denoted N.
```{r}
N = 500 
```
Specify underlying probability of heads.
```{r}
pHeads = 0.5 
```
Flip a coin N times and compute the running proportion of heads at each flip.

Generate a random sample of N flips (heads=1, tails=0):

```{r}
flipSequence = sample( x=c(0,1), prob=c(1-pHeads,pHeads), size=N, replace=TRUE )
```

Compute the running proportion of heads:
```{r}
r = cumsum( flipSequence ) # Cumulative sum: Number of heads at each step.
n = 1:N                    # Number of flips at each step.
runProp = r / n            # Component by component division.
```
Graph the running proportion:
```{r}
plot( n , runProp , type="o" , log="x" , col="skyblue" ,
      xlim=c(1,N) , ylim=c(0.0,1.0) , cex.axis=1.5 ,
      xlab="Flip Number" , ylab="Proportion Heads" , cex.lab=1.5 ,
      main="Running Proportion of Heads" , cex.main=1.5 )
#Plot a dotted horizontal reference line:

abline( h=pHeads , lty="dotted" )
# Display the beginning of the flip sequence:
flipLetters = paste( c("T","H")[flipSequence[1:10]+1] , collapse="" )
displayString = paste0( "Flip Sequence = " , flipLetters , "..." )
text( N , .9 , displayString , adj=c(1,0.5) , cex=1.3 )
# Display the relative frequency at the end of the sequence.
text( N , .8 , paste("End Proportion =",runProp[N]) , adj=c(1,0.5) , cex=1.3 )

```

### Modified Program

Specify the total number of flips, denoted N.
```{r}
N = 500 
```
Specify underlying probability of heads.
```{r}
pHeads = 0.8 
```
Flip a coin N times and compute the running proportion of heads at each flip.

Generate a random sample of N flips (heads=1, tails=0):

```{r}
flipSequence = sample( x=c(0,1), prob=c(1-pHeads,pHeads), size=N, replace=TRUE )
```

Compute the running proportion of heads:
```{r}
r = cumsum( flipSequence ) # Cumulative sum: Number of heads at each step.
n = 1:N                    # Number of flips at each step.
runProp = r / n            # Component by component division.
```
Graph the running proportion:
```{r}
plot( n , runProp , type="o" , log="x" , col="skyblue" ,
      xlim=c(1,N) , ylim=c(0.0,1.0) , cex.axis=1.5 ,
      xlab="Flip Number" , ylab="Proportion Heads" , cex.lab=1.5 ,
      main="Running Proportion of Heads" , cex.main=1.5 )
#Plot a dotted horizontal reference line:

abline( h=pHeads , lty="dotted" )
# Display the beginning of the flip sequence:
flipLetters = paste( c("T","H")[flipSequence[1:10]+1] , collapse="" )
displayString = paste0( "Flip Sequence = " , flipLetters , "..." )
text( N , .9 , displayString , adj=c(1,0.5) , cex=1.3 )
# Display the relative frequency at the end of the sequence.
text( N , .8 , paste("End Proportion =",runProp[N]) , adj=c(1,0.5) , cex=1.3 )

```

## Ex 4.4

Simple Probability Density Function:

Consider Spinner with [0,1] scale on circumference. Biased such that: $p(x)=6x(1-x)$ over the interval $x\in [0,1]$

###Plot this density function

Adapt from IntegralOfDensity.R
```{r}
#source("DBDA2E-utilities.R")
# Graph of normal probability density function, with comb of intervals.
# meanval = 0.0               # Specify mean of distribution.
# sdval = 0.2                 # Specify standard deviation of distribution.
xlow  = 0 # Specify low end of x-axis.
xhigh = 1 # Specify high end of x-axis.
dx = 0.1               # Specify interval width on x-axis
# Specify comb of points along the x axis:
newx = seq( from = xlow , to = xhigh , by = dx )
# Compute y values, i.e., probability density at each value of x:
y = (6*newx)*(1-newx)#( 1/(sdval*sqrt(2*pi)) ) * exp( -.5 * ((x-meanval)/sdval)^2 )
# Plot the function. "plot" draws the intervals. "lines" draws the bell curve.
#openGraph(width=7,height=5)
plot( newx , y , type="h" , lwd=1 , cex.axis=1.5
  , xlab="x" , ylab="p(x)" , cex.lab=1.5 ,
	, main="Normal Probability Density" , cex.main=1.5 )
lines( newx , y , lwd=3 ,  col="skyblue" )
# Approximate the integral as the sum of width * height for each interval.
area = sum( dx * y )
# Display info in the graph.
# text( meanval-sdval , .9*max(y) , bquote( paste(mu ," = " ,.(meanval)) )
#       , adj=c(1,.5) , cex=1.5 )
# text( meanval-sdval , .75*max(y) , bquote( paste(sigma ," = " ,.(sdval)) )
#       , adj=c(1,.5) , cex=1.5 )
# text( meanval+sdval , .9*max(y) , bquote( paste(Delta , "x = " ,.(dx)) )
#       , adj=c(0,.5) , cex=1.5 )
# text( meanval+sdval , .75*max(y) ,
#       bquote(
#         paste( sum(,x,) , " " , Delta , "x p(x) = " , .(signif(area,3)) )
#       ) , adj=c(0,.5) , cex=1.5 )

```

## Ex 4.5 Using normal curve to describe beliefs

```{r}
# Graph of normal probability density function, with comb of intervals.
meanval = 162               # Specify mean of distribution.
sdval = 15               # Specify standard deviation of distribution.
xlow  = meanval - sdval # Specify low end of x-axis.
xhigh = meanval + sdval # Specify high end of x-axis.
dx = sdval/10               # Specify interval width on x-axis
# Specify comb of points along the x axis:
x = seq( from = xlow , to = xhigh , by = dx )
# Compute y values, i.e., probability density at each value of x:
y = ( 1/(sdval*sqrt(2*pi)) ) * exp( -.5 * ((x-meanval)/sdval)^2 )
# Plot the function. "plot" draws the intervals. "lines" draws the bell curve.
#openGraph(width=7,height=5)
plot( x , y , type="h" , lwd=1 , cex.axis=1.5
  , xlab="x" , ylab="p(x)" , cex.lab=1.5 ,
	, main="Normal Probability Density" , cex.main=1.5 )
lines( x , y , lwd=3 ,  col="skyblue" )
# Approximate the integral as the sum of width * height for each interval.
area = sum( dx * y )
# Display info in the graph.
text( meanval-sdval , .9*max(y) , bquote( paste(mu ," = " ,.(meanval)) )
      , adj=c(1,.5) , cex=1.5 )
text( meanval-sdval , .75*max(y) , bquote( paste(sigma ," = " ,.(sdval)) )
      , adj=c(1,.5) , cex=1.5 )
text( meanval+sdval , .9*max(y) , bquote( paste(Delta , "x = " ,.(dx)) )
      , adj=c(0,.5) , cex=1.5 )
text( meanval , .75*max(y) ,
      bquote(
        paste( sum(,x,) , " " , Delta , "x p(x) = " , .(signif(area,3)) )
      ) , adj=c(0,.5) , cex=1.5 )
# Save the plot to an EPS file.
#saveGraph( file = "IntegralOfDensity" , type="eps" )

```



